### 2016-02-08 mcmc.tex
bits

% to-do
% -----
% - Kill all issues in this list, below:
% - Deal with all mentions of ``DFM'' or ``HOGG''
% - Go through github issues, if any.
% - Systematically cite chapters of Brooks \etal in each section.
% - mention ESJD
% - mention Gelman--Rubin statistic
% - mention JAGS and STAN in methods section
% - follow up the idea that burn-in doesn't exist at all!  It does for emcee.
% - need to mention case where there is an MC integral INSIDE the likelihood function
% - need to mention testing the likelihood function, by, say, 1-d scans back and forth.
% - would love to have the Note symbols to be links to the notes and the notes to have links back to the main text...!
% - send to friendlies for comments
%   - Gelman
%   - Goodman
%   - Lang
%   - Rix
%   - Bovy
%   - Brewer
% - post on arXiv

@@

The simplest heuristic proxy statistic for tuning is the \emph{acceptance fraction}.
If you are accepting almost all proposed steps, your step sizes are too small on average.
If you are accepting almost none, your step sizes are too large.
The Goldilocks value is between a half and about a quarter,
  with an argument floating around that it should be 0.234 for best
  performance in high dimensional problems\note{%
    The argument and assumptions underlying the 0.234 fraction
    (and higher acceptance fractions at lower numbers of dimensions)
    are laid out in \citet{gelman234}.}
  (though you could never tune it precisely enough to warrant that third digit of accuracy).
In the burn-in phase (discussed below in \sectionname~\ref{sec:initialization}) of an \MCMC\ run,
  it makes sense to track the acceptance ratio,
  and adjust the proposal distribution variance as you get acceptance ratios
  that are far from the Goldilocks ratio.
This process can be automated easily;
  such automation is part of many projects that use \MCMC.

% HOGG:  Is the following a repeat from previously and/or does it mesh with what's earlier?

A very common---%
  and very useful---%
  kind of proposal distribution is one that cycles through parameters,
  taking a random step in just one parameter at a time.
Aside from involving some persistence of state,
  this kind of proposal distribution can be valuable,
  in part because it reduces the (almost impossible) $D$-dimensional tuning problem
  to $D$ one-dimensional problems:
In this form of proposal distrubution,
  it is possible to track a separate acceptance fraction for every parameter.
Code can be built that uses the burn-in phase to tune all $D$ proposal variances
  such that each of them, individually,
  obtains acceptance at the same Goldilocks ratio.

>>>>

% to-do
% -----
% - Kill all issues in this list, below:
% - Deal with all mentions of ``DFM'' or ``HOGG''
% - Go through github issues, if any.
% - Systematically cite chapters of Brooks \etal in each section.
% - mention Gelman--Rubin statistic
% - mention JAGS and STAN in methods section
% - follow up the idea that burn-in doesn't exist at all!  It does for emcee.
% - need to mention case where there is an MC integral INSIDE the likelihood function
% - need to mention testing the likelihood function, by, say, 1-d scans back and forth.
% - would love to have the Note symbols to be links to the notes and the notes to have links back to the main text...!
% - send to friendlies for comments
%   - Gelman
%   - Goodman
%   - Lang
%   - Rix
%   - Bovy
%   - Brewer
% - post on arXiv

@@

The simplest heuristic proxy statistic for tuning is the \emph{acceptance fraction}.
If you are accepting almost all proposed steps, your step sizes are too small on average.
If you are accepting almost none, your step sizes are too large.
The Goldilocks value is between a half and about a quarter,
  with an argument floating around that it should be 0.234 for best
  performance in high dimensional problems\note{%
    The argument and assumptions underlying the 0.234 fraction
    (and higher acceptance fractions at lower numbers of dimensions)
    are laid out in \citet{gelman234}.\label{note:234}}
  (though you could never tune it precisely enough to warrant that third digit of accuracy).
In the burn-in phase (discussed below in \sectionname~\ref{sec:initialization}) of an \MCMC\ run,
  it makes sense to track the acceptance ratio,
  and adjust the proposal distribution variance as you get acceptance ratios
  that are far from the Goldilocks ratio.
This process can be automated easily;
  such automation is part of many projects that use \MCMC.

A very common---%
  and very useful---%
  kind of proposal distribution is one that cycles through parameters,
  taking a random step in just one parameter at a time.\note{This has a lot to do with Gibbs sampling, which is discussed in \sectionname~\ref{sec:methods}.}
Aside from involving some persistence of state,
  this kind of proposal distribution can be valuable,
  in part because it reduces the (almost impossible) $D$-dimensional tuning problem
  to $D$ one-dimensional problems:
In this form of proposal distrubution,
  it is possible to track a separate acceptance fraction for every parameter.
Code can be built that uses the burn-in phase to tune all $D$ proposal variances
  such that each of them, individually,
  obtains acceptance at the same Goldilocks ratio.
